### 2023-01-27 outputs/paper/paper.qmd
Download raw data and create paper

---
title: "Analysis of Immunization Coverage in Toronto"
author: "Julie Nguyen"
format: pdf
abstract: this is an abstract
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).